Community-based marine reserves in Mexico: do they work?

### 2018-02-27 GainesLabPresentation.Rmd
Added presentation

---
title: "Community-based marine reserves in Mexico: their effectiveness"
subtitle: "(Or their lack of?)"
author: "Juan Carlos Villaseñor-Derbez"
date: "27/02/18"
output: xaringan::moon_reader
---

@@

@@

```{r}
data.frame(Community = c("Isla Natividad", "Maria Elena", "Punta Herrero"),
           N_reserves = c(2, 3, 3),
           N_reserves2 = c("1", "1", "2*"),
           Size = c(10.58, 483.98, 153.13),
           Implemented = c(2006, 2012, 2013)) %>% 
  knitr::kable(format = "html", col.names = c("Community", "Reserves", "Evaluated", "Size (ha)", "Implemented"))
```

@@

- The objectives are *"mixed"* on paper (consrvation + fisheries)
- Ask a fisher, they'll say:

> "We want to increase lobsters populations and catches"

@@

@@

Lobster is their most important resource

- (*Panulirus interruptus* and *P. argus*)

@@

## What are the enabling conditions?

@@

@@

## Fisheries side

@@

- Thanks Ere!

@@

@@

@@

<table style="text-align:center">
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td colspan="6"><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="6" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td colspan="3">Catches (tones)</td><td colspan="3">Revenue(M MXP)</td></tr>
<tr><td style="text-align:left"></td><td>(IN)</td><td>(ME)</td><td>(PH)</td><td>(IN)</td><td>(ME)</td><td>(PH)</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">PostbPost</td><td>7.37</td><td>5.83</td><td>-1.26</td><td><strong>14.37<sup>***</sup></strong></td><td>1.24</td><td>-0.06</td></tr>
<tr><td style="text-align:left">Constant</td><td>122.68<sup>***</sup></td><td>11.41<sup>***</sup></td><td>6.06<sup>***</sup></td><td>10.89<sup>***</sup></td><td>2.04<sup>***</sup></td><td>1.06<sup>***</sup></td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Observations</td><td>15</td><td>14</td><td>13</td><td>15</td><td>14</td><td>13</td></tr>
<tr><td style="text-align:left">R<sup>2</sup></td><td>0.02</td><td>0.04</td><td>0.01</td><td>0.52</td><td>0.04</td><td>0.0004</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td colspan="6" style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>

@@

![](./Img/econ2.png)

@@

![](./Img/econ1.png)

@@

@@

#

>>>>

---
title: "Community-based marine reserves in Mexico: do they work?"
author: "JC"
date: "27/02/18"
output:
  xaringan::moon_reader:
    nature:
      slideNumberFormat: ""
---

@@

# Before we start

- This is onging work

- Other collaborators

- Help me better tell the story (if there is one)

> Even when a community is very well organized, environmental variation can hinder reserve effectiveness

> Even if environment is stable, lack of community organization can hinder reserve effectiveness

@@

```{r}
data.frame(Community = c("Isla Natividad", "Maria Elena", "Punta Herrero"),
           N_reserves = c(2, 3, 3),
           N_reserves2 = c("1", "1", "2*"),
           Size = c(10.58, 9.73, 527.79),
           Implemented = c(2006, 2012, 2013)) %>% 
  knitr::kable(format = "html", col.names = c("Community", "Reserves", "Evaluated", "Size (ha)", "Implemented"))
```

@@

- The objectives are *"mixed"* on paper (conservation + fisheries)

@@

- Ask a fisher, they'll say:

> "We want to increase lobster populations and catches"

@@

@@

## How do environmental and social dynamics influence the outcomes of community-based marine reserves?

@@

Combine:

- Causal inference

- Operationalization of the social-ecological systems framework (Gracias Ere!)

@@

## Fisheries-side

@@

@@

.pull-left[
Some general trends:
- Positive effects in invertebrate abundances
- Fish richness post 2010 takes a dive (IN)
- Negative effect in invertebrate richness (IN)
- Not much more...
]

.pull-right[
![](./Img/plot.png)
]

@@

Are they effective for lobster?

@@

Are they effective for lobster?

- Landings and income

<table style="text-align:center">
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td colspan="6"><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="6" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td colspan="3">Landings (tones)</td><td colspan="3">Revenue(M MXP)</td></tr>
<tr><td style="text-align:left"></td><td>(IN)</td><td>(ME)</td><td>(PH)</td><td>(IN)</td><td>(ME)</td><td>(PH)</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Post</td><td>7.37</td><td>5.83</td><td>-1.26</td><td><strong>14.37<sup>***</sup></strong></td><td>1.24</td><td>-0.06</td></tr>
<tr><td style="text-align:left">Constant</td><td>122.68<sup>***</sup></td><td>11.41<sup>***</sup></td><td>6.06<sup>***</sup></td><td>10.89<sup>***</sup></td><td>2.04<sup>***</sup></td><td>1.06<sup>***</sup></td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Observations</td><td>15</td><td>14</td><td>13</td><td>15</td><td>14</td><td>13</td></tr>
<tr><td style="text-align:left">R<sup>2</sup></td><td>0.02</td><td>0.04</td><td>0.01</td><td>0.52</td><td>0.04</td><td>0.0004</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td colspan="6" style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>

@@

Income

```{r, warning = F, message = F}

load("./Data/conapesca_2000-2015.RData")

cpi <- read.csv("./Data/CPI.csv", stringsAsFactors = F, strip.white = T)

cpi_2014 <- cpi$CPI[cpi$Ano == 2014]

conapesca %<>%
  filter(UnidadEconomica %in% c("Scpp cozumel scl",
                                "Scpp jose maria azcorra scl",
                                "Scpp buzos y pescadores de la baja california scl")) %>% 
  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
  group_by(UnidadEconomica, Ano, NombreComun) %>%
  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
            PesoVivo = sum(PesoVivo),
            Valor = sum(Valor)) %>%
  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
  left_join(cpi, by = "Ano") %>% 
  mutate(Valor = Valor * (CPI / cpi_2014))

L_conapesca <- tibble(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl",
                                          "Scpp cozumel scl",
                                          "Scpp jose maria azcorra scl"),
                      Comunidad = c("Isla Natividad",
                                    "Maria Elena",
                                    "Punta Herrero")) %>% 
  left_join(conapesca, ., by = "UnidadEconomica") %>% 
  mutate(Post = case_when(Comunidad == "Isla Natividad" && Ano <= 2006 ~ 0,
                          Comunidad == "Maria Elena" && Ano <= 2012 ~ 0,
                          Comunidad == "Punta Herrero" && Ano <= 2013 ~ 0,
                          TRUE ~ 1))

L_conapesca2 <- L_conapesca %>% 
  group_by(Comunidad, Post) %>% 
  summarize(meanC = mean(PesoVivo/1000),
            meanI = mean(Valor/1e6))

left_join(L_conapesca, L_conapesca2, by = c("Comunidad", "Post")) %>% 
  ggplot(aes(x = Ano, y = Valor/1e6, color = Comunidad, group = paste(Comunidad, Post), fill = Comunidad)) +
  geom_line(aes(y = meanI), size = 1) +
  geom_point(aes(shape = Comunidad), size = 2) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  geom_vline(xintercept = c(2006.5, 2012.5, 2013.5), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
  labs(x = "Year", y = "Income (M MXP)") +
  theme(legend.position = "none",
        text = element_text(size = 12)) +
  cowplot::theme_cowplot()
```

@@

Landings

```{r}
left_join(L_conapesca, L_conapesca2, by = c("Comunidad", "Post")) %>% 
  ggplot(aes(x = Ano, y = PesoVivo/1e3, color = Comunidad, group = paste(Comunidad, Post), fill = Comunidad)) +
  geom_line(aes(y = meanC), size = 1) +
  geom_point(aes(shape = Comunidad), size = 2) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  geom_vline(xintercept = c(2006.5, 2012.5, 2013.5), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
  labs(x = "Year", y = "Landings (Tones)") +
  theme(legend.position = "none",
        text = element_text(size = 12)) +
  cowplot::theme_cowplot()
```

@@

Isla Natividad

- Designed for invertebrates

- Very well organized community

- Relatively isolated, but still surrounded by "good fishers"

- Reserve is well enforced

- Reserve has been in place quite some time

- Should be the perfect example... but

- Has had mass mortality events due to hypoxia (2009, 2010) and warm water (~2015)

Maria Elena

- Also intended to protect grouper reproductive aggregations

- Well enforced

- Community is very well organized

- No hypoxia or other environmental disturbances

- Small (~ 10 ha)

- Fairly young (5 years of data)

Punta Herrero

- Also intended to protect grouper reproductive aggregations

- Good size (~ 500 ha)

- Low enforcement

- Community is not very well organized

- No hypoxia or other environmental disturbances

- Fairly young (4 years of data)

# Telling the story

> Even when a community is very well organized, environmental variation can hinder reserve effectiveness

> Even if environment is stable, lack of community organization can hinder reserve effectiveness

## Missing pieces

- Data on environmental disturbances (can get DO, S, T > 2009 for N)

---
background-image: url(https://raw.githubusercontent.com/jcvdav/ReserveEffect/master/Img/Imagen5.jpg)
class: inverse, center, middle
background-size: cover

# Thanks!

## Questions?

## Suggestions?

## Comments?

### Anyone??

# Help!

# Extras

@@
